\documentclass{article}

\begin{document}

\begingroup
\centering
\Large
\textbf{Welcome to the 39th CCSC Southeastern Conference}\par
\endgroup

\vspace{10pt}

Welcome to the 39th Southeastern Regional Conference of the Consortium
for Computing Sciences in Colleges. The CCSC:SE Regional Board welcomes
you to Macon, GA, the home of Mercer University. The conference promotes
a productive exchange of information among college personnel concerned
with producing quality computer-oriented curricula as well as using
effective educational methods to teach computer science. It is for
faculty as well as administrators of academic computing facilities, and
for students to participate in activities that promote computer
science. We hope that you will find something to challenge yourself and
engage you at the conference!

The robust conference program highlighted by four sessions each with
multiple tracks, including engaging guest speakers, workshops, a panel
discussion, student posters, a nifty assignment session and four
sessions of high-quality refereed papers. We received fifty-seven papers
this year and accepted eighteen papers for the conference and included
in the proceedings – an acceptance rate of 32%.

Two exciting activities designed specifically for students – a research
contest and an undergraduate programming competition, with prizes for
the top finishers in each.

We especially would like to thank the faculty, staff, and students of
Mercer for their help in organizing and publicizing this conference. We
also extend heartfelt thanks to the CCSC Board, the CCSC:SE Regional
Board, and to a wonderful Conference Committee, led by Conference Chair
and programming contest coordinator Dr. Andy Digh, and research contest
coordinator Dr. Fahad Sultan. Thank you all so much for your time and
energy.

We also need to send our deepest appreciation to our partners, sponsors,
and vendors. Please take the time to go up to them and thank them for
their contributions and support for computing sciences education -- CCSC
National Partners: Rephactor, ACM2y, and Blossoms. Sponsoring
Organizations: CCSC, ACM-SIGCSE, and Middle Georgia STEM Alliance and
21st Century.

We could not have done this without the excellent submissions from
authors, the insightful comments from volunteer reviewers, and the
support from our editor Dr. Adam Lewis. Thanks to all of you for helping
to create such a strong program for this year’s conference.

We hope you enjoy the conference and your visit to Mercer.

\vspace{10pt}

\hfill Karen Works

\hfill Florida State University

\hfill Regional Chair

\vspace{10pt}

\hfill Andy Digh

\hfill Mercer University

\hfill Conference Chair and Host

\end{document}
