\documentclass{article}

\input{preamble}

\title{Best Practices for Face-to-face and Online Undergraduate Research\\
\vspace{0.2in}
\large{
Conference Tutorial
}}

\author{
Karen E. Works\\
Computer Science\\
Florida State University\\
Panama City, FL 32405\\
\email{keworks@fsu.edu}
}

\begin{document}
\maketitle

\section*{Abstract}
Undergraduate research is beneficial to students in so many ways. Crowe and Burke \cite{crowe2008assessing} found that it increases the number of students who go to graduate school and that students who participate report a higher level of satisfaction with their undergraduate experience. It has been shown to increase student engagement \cite{fechheimer2011well}. Retention in computing majors \cite{redfield2024longitudinal} and increase in problem solving skills \cite{seifan2022effect} have been correlated with students who participate in undergraduate research. Given all the benefits to students, how do you get started with undergraduate research at your college?

Dr. Karen Works has co-chaired the student research symposium at Florida State University on the Panama City campus since Fall 2020. The symposium supports face-to-face poster sessions, online pre-recorded presentations, and online video conferencing presentations. In addition, she has been a research mentor for undergraduate computer science students outside of her courses, both online and face-to-face.

In this tutorial, she will share her experience in coordinating and hosting a student research symposium on the FSU Panama City campus discussing both the challenges and rewards in working with students on undergraduate research projects, to encourage others to become research mentors. Best practices to support undergraduate research in different modalities, namely face-to-face and on-line, will be presented. This tutorial session will introduce how instructors can get undergraduate research started with their students and how to coordinate research symposiums on their campuses. The session will be interactive with discussion activities.  


%\section*{Biography}
%Dr. Works joined Florida State University (FSU) Panama City in the summer of 2019. She teaches courses for the distant learning computer science degree program. Prior to coming to FSU, she had 6 years university teaching experience in traditional face to face classroom settings.  


\medskip

\bibliographystyle{plain}
\small 
\bibliography{tutorial_abstract}

\end{document}
